**OptiClust: Improved method for assigning amplicon-based sequence data to operational taxonomic units**

### 2016-12-23 submission/Westcott_OptiClust_mSphere_2016.Rmd
Fix missing reference

Several metrics have emerged for assessing the quality of OTU assignment algorithms. These have included the time and memory required to run the algorithm [@Sun2009; @Cai2011; @Rideout2014; @Mah2014], agreement between OTU assignments and the sequences' taxonomy [@Edgar2013; @Mah2014; @Mah2015; @Barriuso2011; @Bonder2012; @Chen2013; @Huse2010; @May2014; @Cai2011; @Sun2011; @White2010; @AlGhalith2016], sensitivity of an algorithm to stochastic processes [@He2015], the number of OTUs generated by the algorithm [@Edgar2013; @Kopylova2016], and the ability to regenerate the assignments made by other algorithms [@Schmidt2014; @Rideout2014]. Unfortunately, these methods fail to directly quantify the quality of the OTU assignments. An algorithm may complete with minimal time and memory requirements or generate an idealized number of OTUs, but the composition of the OTUs could be incorrect. These metrics also tend to be subjective. For instance, a method may appear to be recapitulate the taxonomy of a synthetic community with known taxonomic structure, but do a poor job when applied to real communities with poorly defined taxonomic structure or for sequences that are prone to misclassification. As an alternative, we developed an approach to objectively benchmark the clustering quality of OTU assignments [@Schloss2016a; @Westcott2015; @Schloss2011]. This approach counts the number of true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN) based on the pairwise distances. Sequence pairs that are within the user-specified threshold and are clustered together represent TPs and those in different OTUs are FNs. Those sequence pairs that have a distance larger than the threshold and are not clustered in the same OTU are TNs and those in the same OTU are FPs. These values can be synthesized into a single correlation coefficient, the Matthews' Correlation Coefficient (MCC), which measures the correlation between observed and predicted classifications and is robust to cases where there is an uneven distribution across the confusion matrix [@Matthews1975; @Powers2011]. Consistently, the average neighbor algorithm was identified as among the best or the best algorithm. The distance-based greedy clustering as implemented in VSEARCH has also performed well. The computational resources required to complete the average neighbor algorithm can be significant for large datasets and so there is a need for an algorithm that efficiently produces consistently high quality OTU assignments.

>>>>

Several metrics have emerged for assessing the quality of OTU assignment algorithms. These have included the time and memory required to run the algorithm [@Sun2009; @Cai2011; @Rideout2014; @Mah2014], agreement between OTU assignments and the sequences' taxonomy [@Edgar2013; @Mah2014; @Mah2015; @Barriuso2011; @Bonder2012; @Chen2013; @Huse2010; @May2014; @Cai2011; @Sun2011; @White2010; @AlGhalith2016], sensitivity of an algorithm to stochastic processes [@He2015], the number of OTUs generated by the algorithm [@Edgar2013; @Kopylova2016], and the ability to regenerate the assignments made by other algorithms [@Schmidt2014; @Rideout2014]. Unfortunately, these methods fail to directly quantify the quality of the OTU assignments. An algorithm may complete with minimal time and memory requirements or generate an idealized number of OTUs, but the composition of the OTUs could be incorrect. These metrics also tend to be subjective. For instance, a method may appear to be recapitulate the taxonomy of a synthetic community with known taxonomic structure, but do a poor job when applied to real communities with poorly defined taxonomic structure or for sequences that are prone to misclassification. As an alternative, we developed an approach to objectively benchmark the clustering quality of OTU assignments [@Schloss2016a; @Westcott2015; @Schloss2011]. This approach counts the number of true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN) based on the pairwise distances. Sequence pairs that are within the user-specified threshold and are clustered together represent TPs and those in different OTUs are FNs. Those sequence pairs that have a distance larger than the threshold and are not clustered in the same OTU are TNs and those in the same OTU are FPs. These values can be synthesized into a single correlation coefficient, the Matthews' Correlation Coefficient (MCC), which measures the correlation between observed and predicted classifications and is robust to cases where there is an uneven distribution across the confusion matrix [@Matthews1975]. Consistently, the average neighbor algorithm was identified as among the best or the best algorithm. The distance-based greedy clustering as implemented in VSEARCH has also performed well. The computational resources required to complete the average neighbor algorithm can be significant for large datasets and so there is a need for an algorithm that efficiently produces consistently high quality OTU assignments.
